---
title: "Reshape_HMD"
author: "Jessica McCordic"
date: "2025-08-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load libraries

```{r load lib}

library(PAMscapes)
library(PAMmisc)
library(tidyverse)

```

Load HMD data

```{r load data}

# bring in hourly HMD spectra

dir_hmd <- "data/HMD_hourly"

hmd_tables <- dir_hmd |>
    # list all files with .csv extension
    list.files(pattern = ".csv") |>
    map_chr(~paste0(dir_hmd,"\\",  .)) |>
    # use map to iterate read.delim() function over each file in the directory
    map(~read_csv(.))

all_hmd_og <- do.call("rbind", hmd_tables) 

```


Clean up table a bit and add columns for grouping

```{r add columns}

hmd_data <- all_hmd_og |>
  # remove NA values from data gaps or start/end of deployment
  filter(!is.na(HMD_10)) |> 
  # add grouping variables for dates
  mutate(hour_UTC = hour(UTC),
         day_UTC = day(UTC),
         month_UTC = month(UTC),
         year_UTC = year(UTC),
          # create NPZ column to simplify plot labels
    npz_id = str_replace(platform, "Parks Australia Cod Grounds Marine Park", "Cod Grounds"),
    npz_id = str_replace(npz_id, "Parks Australia Solitary Islands Marine Park, West Pimpernel mooring", "Solitary Islands"),
    npz_id = str_replace(npz_id, "Parks Australia Dampier Marine Park", "Dampier"),
    npz_id = str_replace(npz_id, "Parks Australia Ningaloo Marine Park, North mooring", "Ningaloo"),
    npz_id = str_replace(npz_id, "Parks Australia Jurien Marine Park, Northeast mooring", "Jurien"),
    npz_id = str_replace(npz_id, "Parks Australia Geographe Marine Park", "Geographe"),
    npz_id = str_replace(npz_id, "Parks Australia Two Rocks Marine Park, West mooring", "Two Rocks"),
    npz_id = str_replace(npz_id, "Parks Australia South-west Corner Marine Park, South mooring", "South-west Corner"),
    npz_id = str_replace(npz_id, "Parks Australia Murat Marine Park, East mooring", "Murat"),
    
    # add network for grouping/faceting later
    network = case_when(platform == "Parks Australia Cod Grounds Marine Park" | 
                          platform == "Parks Australia Solitary Islands Marine Park, West Pimpernel mooring" 
                        ~ "Temperate East",
                        platform == "Parks Australia Dampier Marine Park" | 
                          platform == "Parks Australia Ningaloo Marine Park, North mooring" ~ "Northwest",
                        TRUE ~ "South-west"
    ),
    network = fct(network),
    network = fct_relevel(network, c("Temperate East","Northwest"))) |>
  
   rowwise() |> 
  # assign categories based on sun elev
  mutate(sun_elev_rad = suncalc::getSunlightPosition(UTC, 
                                                     Latitude,
                                                     Longitude,
                                                     keep = "altitude")$altitude,
         sun_elev_deg = (sun_elev_rad*180)/(pi),
         light_reg = case_when(
           # if sun elev is positive, daytime
           sun_elev_deg > 0 ~ "day",
           # if sun elevation is less than -12, nighttime
           sun_elev_deg < -12 ~ "night",
           # if between 0 and -12, "twilight"
           sun_elev_deg <= 0 & sun_elev_deg >= -12 ~ "twilight"
         )
  ) 

```

```{r LTSA}


# subset to pull 1 representative month from each deployment

ltsa_df <- all_hmd_data |>
  group_by(Dep_ID) |>
  mutate(month_dep = month(UTC), 
         ltsa_keep = case_when(
           # select months to use for LTSA
           Dep_ID == "PARKSAUSTRALIA_CGMP_201807_CG" ~ 7,
           Dep_ID == "PARKSAUSTRALIA_CGMP_201812_CG" ~ 1,
           Dep_ID == "PARKSAUSTRALIA_CGMP_201904_CG" ~ 5,     
           Dep_ID == "PARKSAUSTRALIA_DAMPIER_202009_DNE" ~ 10,  
           Dep_ID == "PARKSAUSTRALIA_MURAT_202002_MRE" ~ 4,     
           Dep_ID == "PARKSAUSTRALIA_NINGALOO_201909_NGN" ~ 10,
           Dep_ID == "PARKSAUSTRALIA_SIMP_201808_WP" ~ 8, 
           Dep_ID == "PARKSAUSTRALIA_SIMP_201812_WP" ~ 12,       
           Dep_ID == "PARKSAUSTRALIA_SIMP_201906_WP" ~ 7,     
           Dep_ID == "PARKSAUSTRALIA_SWC_202201_SWS" ~ 2)
         ) |>
  filter(month_dep == ltsa_keep) |>
  mutate(ltsa_label = paste0(npz_id, ": ", month.name[month_dep]))

year(ltsa_df$UTC) <- 2019
# 
# plot_sep_LTSA <- function(ltsa_lab, ltsa_df) {
#   ltsa_sub <- ltsa_df |>
#     filter(ltsa_label == ltsa_lab)
#   
#   plotLTSA(x = ltsa_sub,
#             bin = '12hour',
#          scale = "log",
#          units= expression(paste("dB re 1 \u00b5Pa" ^ "2", "/Hz")))+
#     theme_bw() +
#   theme(axis.text.x = element_text(size= 18),
#         axis.text.y= element_text(size= 18),
#         legend.title= element_text(size= 18),
#         legend.text= element_text(size= 18))
# 
# }
# 
# plot_sep_LTSA(ltsa_lab = ltsa_df$ltsa_label[1], ltsa_df = ltsa_df)
# 
ltsa_plot_data <- plotLTSA(x=ltsa_df, 
         bin = '4hour',
         scale = "log",
         units= expression(paste("dB re 1 \u00b5Pa" ^ "2", "/Hz")), 
         facet = "ltsa_label",
         maxBins = 2500) +
  facet_wrap(vars(ltsa_label), nrow = 10,
             scales = "free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(size= 18),
        axis.text.y= element_text(size= 18),
        strip.text= element_text(size= 18),
        legend.title= element_text(size= 18),
        legend.text= element_text(size= 18),
        plot.margin=margin(0, 0, 0, 1,unit = "cm"))

ggsave(plot = ltsa_plot_data, 
       filename = paste0("figs/", "Monthly_rep_LTSA_4h.png"), width=12, height=20,
        units="in", dpi=300)


```


```{r month grouping}

# group to month level  

hmd_data_month <- hmd_data |>
  # group to day level per deployment
  group_by(Dep_ID, year_UTC, month_UTC, day_UTC) |>
  # only keep full days
  filter(n_distinct(hour_UTC) == 24) |>
  
  # ungroup and regroup back out to month level per deployment
  ungroup() |>
  group_by(network, npz_id, Dep_ID, year_UTC, month_UTC) |>
  
  # count days of effort per month
  mutate(effort_days = n_distinct(day_UTC))

```


Get median PSD per month per deployment

```{r median PSD}

psd_month <- hmd_data_month |>
  # group by month, include effort_days in grouping so it's retained in final df
  group_by(network, npz_id, Dep_ID, year_UTC, month_UTC, effort_days) |>  
  # use plotPSD with returnData = TRUE to get values of monthly PSDs
  # use group_modify to calculate per group
  group_modify(~plotPSD((.), 
                        style = "quantile", 
                        q = c(0.1, 0.5, 0.9), 
                        returnData = TRUE))


write_csv(psd_month, "data/Df_analysis/All_deps_monthly_PSD.csv")



```


Get median PSD per diel period per month per deployment

```{r median PSD}

psd_month_light <- hmd_data_month |>
  # group by week, include effort_days in grouping so it's retained in final df
  group_by(network, npz_id, Dep_ID, month_UTC, effort_days, light_reg) |>  
  # use plotPSD with returnData = TRUE to get values 
  # use group_modify to calculate per group
  group_modify(~plotPSD((.), 
                        style = "quantile", 
                        q = c(0.1, 0.5, 0.9), 
                        returnData = TRUE))

write_csv(psd_month_light, "data/Df_analysis/All_deps_monthly_diel_PSD.csv")
